---
title: "Reproducible Research: Peer Assessment 1"
output: 
    html_document: 
        keep_md: true
---

```{r,echo=FALSE}
library(data.table)
library(mice)
library(ggplot2)
library(DMwR)
```


## Loading and preprocessing the data
```{r, echo=TRUE}
if(!file.exists("./data.zip")){
    download.file(url, "data.zip")
    unzip("data.zip",exdir=".")
}
df<- read.csv("activity.csv")
head(df)
dim(df)
```
## What is mean total number of steps taken per day?
```{r, echo=TRUE}
total_steps_per_day <- aggregate(steps ~ date, df, sum)
hist(total_steps_per_day$steps, xlab = "steps", main="Frequency of steps")
mean_steps <- mean(total_steps_per_day$steps, na.rm = TRUE)
median_steps <- median(total_steps_per_day$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?
```{r, echo=TRUE}
ggplot(total_steps_per_day, aes(as.Date(date, format = "%Y-%m-%d"), steps)) + geom_line() + xlab("") + ylab("Daily avg steps") + scale_x_date(date_breaks = "1 weeks", date_labels = "%b %d")
```
Average number of steps by interval and the interval which has maximum average steps
```{r, echo=TRUE}
by_interval <- aggregate(steps ~ interval, df, mean)
max_steps_interval <- by_interval[which.max(by_interval$steps),]

```

## Imputing missing values
Showing number of missing values
```{r, echo=TRUE}
md.pattern(df, plot=TRUE)
```
Imputing missing values using knn imputation
```{r, echo=TRUE}
imputedData <- knnImputation(df)
```
```{r,echo=TRUE}
total_steps_imputed <- aggregate(steps ~ date, imputedData, sum)
hist(total_steps_imputed$steps, col="red",main="Frequency of steps")
hist(total_steps_per_day$steps, xlab = "steps",col="blue", ylim=c(0,40),add = T)

legend("topright", c("imputed", "missing"), fill=c("red", "blue"))
box()
```


## Are there differences in activity patterns between weekdays and weekends? 
```{r, echo=TRUE}
imputedData$week <- ifelse(!(weekdays(as.Date(imputedData$date)) %in% c("Saturday", "Sunday")),"weekday","weekend")
total_steps_interval <- aggregate(steps ~ interval+ week, imputedData, mean)
ggplot(total_steps_interval, aes(interval,steps)) + geom_line() + facet_grid(week ~ .)
```
